---
title: "Local Moran's I with p-value adjustment"
date: "`r Sys.Date()`"
author: "Sven Lautenbach"
output:
  rmdformats::readthedown:
      rmdformats::readthedown:
      self_contained: true
      thumbnails: true
      lightbox: true
      gallery: false
      highlight: tango
---

```{r setup}
library(spdep)
library(sf)
library(geojsonsf)
library(tmap)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(lwgeom)
```

# Preprocessing


```{r}
are <- geojson_sf("../data/are_mean_median_kontur_corpo_non_corpo_combi_withMeta.geojson")

idn <- geojson_sf("../data/idn_mean_median_kontur_corpo_non_corpo_combi_withMeta_v2.geojson")
col <- geojson_sf("../data/col_mean_median_kontur_corpo_non_corpo_combi_with_meta_v2.geojson")

st_crs(are) <- 3857
are <- are |> filter(!is.na(fid))

st_crs(idn) <- 3857
idn <- idn |> filter(!is.na(fid))

st_crs(col) <- 3857
col <- col |> filter(!is.na(fid))

```

```{r}
tm_shape(are) + tm_polygons(col="diff_non_corpo")
```


```{r}
library(rgeoboundaries) # to install, execute: remotes::install_github("wmgeolab/rgeoboundaries")
countries <- gb_adm0() # download all countries
countries <- st_transform(countries, crs= st_crs(are))

#countries$shapeName
are_poly <- countries |> filter(shapeName=="United Arab Emirates")
idn_poly <- countries |> filter(shapeName=="Indonesia")
col_poly <- countries |> filter(shapeName=="Colombia")
```

# NB definition

Take care of islands!


```{r}
are_nbQueen <- poly2nb(are, queen= TRUE)
idn_nbQueen <- poly2nb(idn, queen= TRUE)
col_nbQueen <- poly2nb(col, queen= TRUE)

p1 <- as.data.frame(table(card(are_nbQueen))) %>%
  ggplot(aes(x=Var1, y=Freq)) +
  geom_col() + xlab("Number of neighbors") +
  labs(title="United Arab Emirates")
p2 <- as.data.frame(table(card(idn_nbQueen))) %>%
  ggplot(aes(x=Var1, y=Freq)) +
  geom_col() + xlab("Number of neighbors") +
  labs(title="Indonesia")
p3 <- as.data.frame(table(card(col_nbQueen))) %>%
  ggplot(aes(x=Var1, y=Freq)) +
  geom_col() + xlab("Number of neighbors") +
  labs(title="Colombia")

ggarrange(p1, p2, p3)
```

Row standardization

```{r}
are_lwQueen <- nb2listw(are_nbQueen, style= "W", zero.policy = TRUE)
idn_lwQueen <- nb2listw(idn_nbQueen, style= "W", zero.policy = TRUE)
col_lwQueen <- nb2listw(col_nbQueen, style= "W", zero.policy = TRUE)
```

```{r}
plot(st_geometry(are), border = "grey60", reset = FALSE)
plot(are_lwQueen, coords=st_coordinates(st_centroid(are)), col="red",  cex=.7, add=TRUE)
```

# Local Moran's I

```{r function local Morans I }
getLocalMoranFactor <- function(x, listw, pval, quadr = "mean", p.adjust.method ="holm", method="normal", nb)
{
  if(! (quadr %in% c("mean", "median", "pysal")))
    stop("getLocalMoranFactor: quadr needs to be one of the following values: 'mean', 'median', 'pysal'")
  if(! (method %in% c("normal", "perm")))
    stop("getLocalMoranFactor: method needs to be one of the following values: 'normal', 'perm'")
  if(! (p.adjust.method %in% p.adjust.methods))
    stop(paste("getLocalMoranFactor: p.adjust.method needs to be one of the following values:", p.adjust.methods))
  
  # adjust for multiple testing
  if(method == "normal")
    lMc <- spdep::localmoran(x, listw= listw) 
  else
    lMc <- spdep::localmoran_perm(x, listw= listw)
  lMc[,5] <-  p.adjustSP(lMc[,5], method = p.adjust.method, nb = nb)
  lMcQuadr <- attr(lMc, "quadr")
  
  lMcFac <- as.character(lMcQuadr[, quadr])
  # which values are significant
  idx <- which(lMc[,5]> pval)
  lMcFac[idx] <- "Not sign."
  lMcFac <- factor(lMcFac, levels = c("Not sign.", "Low-Low", "Low-High", "High-Low",  "High-High"))
  
  # set those with empty neighbors to NA
  
  idx <- which(card(nb)==0)
  lMcFac[idx] <- "Not sign."
  
  return(list(lMcFac, lMc[,5]))
}
```

## UAE

### diff_non_corpo

```{r}
are_lisa <- getLocalMoranFactor(x= are$diff_non_corpo,
                    listw= are_lwQueen,
                    nb= are_nbQueen,
                    pval=.05)
# local Moran's I quadrants
are$localM_non_corpo <- are_lisa[[1]]
xtabs(~localM_non_corpo, data=are)
# adjusted p-values
are$localM_adjPval_non_corpo <- are_lisa[[2]]
ggplot(are |> filter(localM_adjPval_non_corpo < 0.1), aes(x=localM_adjPval_non_corpo)) +
  geom_histogram(binwidth=0.005) + xlab("Adjusted p-values") +
  labs(title="Histogram for UAE, diff non corpo", subtitle="Only adjusted p-values < 0.1")
```

```{r}
localMcPalette <- c("white", "midnightblue", "lightblue", "lightpink", "red")

#tmap_mode("view")

m_are_non_corpo <- tm_shape(are_poly) + tm_polygons(col="lightgrey") +
  tm_shape(countries) + tm_polygons(col="darkgrey") +
  tm_shape(are) + 
  tm_polygons(col="localM_non_corpo", palette= localMcPalette, 
              legend.hist = FALSE, legend.reverse = TRUE, title = "LISA") +
  tm_layout(legend.outside = FALSE, bg.color = "steelblue", outer.bg.color = "lightgrey",
            legend.outside.size = 0.5, attr.outside = TRUE, legend.bg.color = "white",
            legend.outside.position = "right", main.title = "UAE, diff non corpo",
            outer.margins= c(0,0.02,0,0.02), asp=0) + tm_scale_bar()
```

### Diff corpo

```{r}
are_lisa_corpo <- getLocalMoranFactor(x= are$diff_corpo,
                    listw= are_lwQueen,
                    nb= are_nbQueen,
                    pval=.05)
# local Moran's I quadrants
are$localM_corpo <- are_lisa_corpo[[1]]
xtabs(~localM_corpo, data=are)
# adjusted p-values
are$localM_adjPval_corpo <- are_lisa_corpo[[2]]
ggplot(are |> filter(localM_adjPval_corpo < 0.1), aes(x=localM_adjPval_corpo)) +
  geom_histogram(binwidth=0.005) + xlab("Adjusted p-values") +
  labs(title="Histogram for UAE, diff corpo", subtitle="Only adjusted p-values < 0.1")
```

```{r}
#tmap_mode("view")

m_are_corpo <- tm_shape(are_poly) + tm_polygons(col="lightgrey") +
  tm_shape(countries) + tm_polygons(col="grey") + tm_shape(are) + 
  tm_polygons(col="localM_corpo", palette= localMcPalette, 
              legend.hist = FALSE, legend.reverse = TRUE, title = "LISA") +
  tm_layout(legend.outside = FALSE, bg.color = "steelblue", outer.bg.color = "lightgrey",
            legend.outside.size = 0.5, attr.outside = TRUE, legend.bg.color = "white",
            legend.outside.position = "right", main.title = "UAE, diff corpo",
            outer.margins= c(0,0.02,0,0.02), asp=0) + tm_scale_bar()
```

```{r, fig.height=5, fig.width=10}
tmap_arrange(m_are_non_corpo, m_are_corpo)
```


## Indonesia

### Diff non corpo

```{r}
idn_lisa <- getLocalMoranFactor(x= idn$diff_non_corpo,
                    listw= idn_lwQueen,
                    nb= idn_nbQueen,
                    pval=.05)
# local Moran's I quadrants
idn$localM_non_corpo <- idn_lisa[[1]]
xtabs(~localM_non_corpo, data=idn)
# adjusted p-values
idn$localM_adjPval_non_corpo <- idn_lisa[[2]]

ggplot(idn |> filter(localM_adjPval_non_corpo < 0.1), aes(x=localM_adjPval_non_corpo)) +
  geom_histogram(binwidth=0.005) + xlab("Adjusted p-values") +
  labs("Histogram for Indonesia", subtitle="Only adjusted p-values < 0.1")
```

```{r, fig.height=7, fig.width=11}
m_idn_non_corpo <- tm_shape(idn_poly) + tm_polygons(col="lightgrey") +
  tm_shape(countries) + tm_polygons(col="darkgrey") + 
  tm_shape(idn) + 
  tm_fill(col="localM_non_corpo", palette= localMcPalette, 
              legend.hist = FALSE, legend.reverse = TRUE, title = "LISA",
              border.alpha=.5) +
  tm_layout(legend.outside = FALSE, bg.color = "steelblue", outer.bg.color = "lightgrey",
            legend.outside.size = 0.5, attr.outside = TRUE, legend.bg.color = "white",
            legend.position =c("right", "top"),
            legend.outside.position = "right", main.title = "Indonesia, diff non corpo",
            outer.margins= c(0,0.02,0,0.02), asp=0) + tm_scale_bar()
```

### Diff corpo

```{r}
idn_lisa_corpo <- getLocalMoranFactor(x= idn$diff_corpo,
                    listw= idn_lwQueen,
                    nb= idn_nbQueen,
                    pval=.05)
# local Moran's I quadrants
idn$localM_corpo <- idn_lisa_corpo[[1]]
xtabs(~localM_corpo, data=idn)
# adjusted p-values
idn$localM_adjPval_corpo <- idn_lisa_corpo[[2]]
ggplot(idn |> filter(localM_adjPval_corpo < 0.1), aes(x=localM_adjPval_corpo)) +
  geom_histogram(binwidth=0.005) + xlab("Adjusted p-values") +
  labs(title="Histogram for Indonesia, diff corpo", subtitle="Only adjusted p-values < 0.1")
```

```{r, fig.height=7, fig.width=11}
#tmap_mode("view")

m_idn_corpo <- tm_shape(idn_poly) + tm_polygons(col="lightgrey") +
  tm_shape(countries) + tm_polygons(col="darkgrey") + 
  tm_shape(idn) + 
  tm_fill(col="localM_corpo", palette= localMcPalette, 
              legend.hist = FALSE, legend.reverse = TRUE, title = "LISA") +
  tm_layout(legend.outside = FALSE, bg.color = "steelblue", outer.bg.color = "lightgrey",
            legend.outside.size = 0.5, attr.outside = TRUE, legend.bg.color = "white",
            legend.position =c("right", "top"),
            legend.outside.position = "right", main.title = "Indonesia, diff corpo",
            outer.margins= c(0,0.02,0,0.02), asp=0) + tm_scale_bar()
```

```{r, fig.width=11, fig.height=10}
tmap_arrange(m_idn_non_corpo, m_idn_corpo, nrow = 2)
```


## Colombia

## Diff non corpo

```{r}
col_lisa <- getLocalMoranFactor(x= col$diff_non_corpo,
                    listw= col_lwQueen,
                    nb= col_nbQueen,
                    pval=.05)
# local Moran's I quadrants
col$localM_non_corpo <- col_lisa[[1]]
xtabs(~localM_non_corpo, data=col)
# adjusted p-values
col$localM_adjPval_non_corpo <- col_lisa[[2]]

ggplot(col |> filter(localM_adjPval_non_corpo < 0.1), aes(x=localM_adjPval_non_corpo)) +
  geom_histogram(binwidth=0.005) + xlab("Adjusted p-values") +
  labs("Histogram for Colombia", subtitle="Only adjusted p-values < 0.1")
```

```{r, fig.height=7, fig.width=6}
m_col_non_corpo <- tm_shape(col_poly) + tm_polygons(col="lightgrey") +
  tm_shape(countries) + tm_polygons(col="darkgrey") +
  tm_shape(col) + 
  tm_fill(col="localM_non_corpo", palette= localMcPalette, 
              legend.hist = FALSE, legend.reverse = TRUE, title = "LISA") +
  tm_layout(legend.outside = FALSE, bg.color = "steelblue", outer.bg.color = "lightgrey",
            legend.outside.size = 0.5, attr.outside = TRUE, legend.bg.color = "white",
            legend.outside.position = "right", main.title = "Colombia, diff non corpo",
            outer.margins= c(0,0.02,0,0.02), asp=0) + tm_scale_bar()
```


### Diff corpo

```{r}
col_lisa_corpo <- getLocalMoranFactor(x= col$diff_corpo,
                    listw= col_lwQueen,
                    nb= col_nbQueen,
                    pval=.05)
# local Moran's I quadrants
col$localM_corpo <- col_lisa_corpo[[1]]
xtabs(~localM_corpo, data=col)
# adjusted p-values
col$localM_adjPval_corpo <- col_lisa_corpo[[2]]
ggplot(col |> filter(localM_adjPval_corpo < 0.1), aes(x=localM_adjPval_corpo)) +
  geom_histogram(binwidth=0.005) + xlab("Adjusted p-values") +
  labs(title="Histogram for Colombia, diff corpo", subtitle="Only adjusted p-values < 0.1")
```

```{r, fig.height=7, fig.width=6}
#tmap_mode("view")

m_col_corpo <- tm_shape(col_poly) + tm_polygons(col="lightgrey") +
  tm_shape(countries) + tm_polygons(col="darkgrey") +
  tm_shape(col) + 
  tm_fill(col="localM_corpo", palette= localMcPalette, 
              legend.hist = FALSE, legend.reverse = TRUE, title = "LISA") +
  tm_layout(legend.outside = FALSE, bg.color = "steelblue", outer.bg.color = "lightgrey",
            legend.outside.size = 0.5, attr.outside = TRUE, legend.bg.color = "white",
            legend.outside.position = "right", main.title = "Colombia, diff corpo",
            outer.margins= c(0,0.02,0,0.02), asp=0) + tm_scale_bar()
```

```{r, fig.height=7, fig.width=10}
tmap_arrange(m_col_non_corpo, m_col_corpo )
```

# Confusion tables for clusters between corporate and non corporate mapping activities

## UAE

Absolute number of hexagon cells

```{r}
xtabs(~ localM_non_corpo + localM_corpo, 
      data=are)
```

Percentage

```{r}
xtabs(~ localM_non_corpo + localM_corpo, 
      data=are) |>
  prop.table() |>
  round(4) * 100
```

Raw differences for corporate and non corporate mappers per hexagon cell

```{r}
ggplot(are, aes(x=diff_corpo, y= diff_non_corpo)) +
  geom_point(alpha=.2) +
  xlab("Difference in corporate mapping activity") +
  ylab("Difference in non-corporate mapping activity") +
  labs(title="UAE")
```

## Indonesia

```{r}
xtabs(~ localM_non_corpo + localM_corpo, 
      data=idn)
```

The rows represent the local Moran's I categories for difference of the non corporate mappers while the columns represent the same for the corporate mappers.

To double check:

```{r}
xtabs(~ localM_non_corpo, data=idn)
xtabs(~ localM_non_corpo + localM_corpo, 
      data=idn) |> rowSums()
```


```{r}
xtabs(~ localM_non_corpo + localM_corpo, 
      data=idn)|>
  prop.table() |>
  round(4)*100
```

Raw differences for corporate and non corporate mappers per hexagon cell

```{r}
ggplot(idn, aes(x=diff_corpo, y= diff_non_corpo)) +
  geom_point(alpha=.2) +
  xlab("Difference in corporate mapping activity") +
  ylab("Difference in non-corporate mapping activity") +
  labs(title="Indonesia")
```



## Colombia

```{r}
xtabs(~ localM_non_corpo + localM_corpo, 
      data=col)
```


```{r}
xtabs(~ localM_non_corpo + localM_corpo, 
      data=col)|>
  prop.table() |>
  round(4)*100
```

Raw differences for corporate and non corporate mappers per hexagon cell

```{r}
ggplot(col, aes(x=diff_corpo, y= diff_non_corpo)) +
  geom_point(alpha=.2) +
  xlab("Difference in corporate mapping activity") +
  ylab("Difference in non-corporate mapping activity") +
  labs(title="Columbia")
```

# Write results to geopackage

Currently not evaluated to avoid unnecessary storage at gitlab.

```{r, eval=FALSE}
st_write(dplyr::select(are, !fid), dsn="lisa_diff_non_corpo_corpo.gpkg", layer="are", append=FALSE)
st_write(dplyr::select(idn, !fid), dsn="lisa_diff_non_corpo_corpo.gpkg", layer="idn", append=FALSE)
st_write(dplyr::select(col, !fid), dsn="lisa_diff_non_corpo_corpo.gpkg", layer="col", append=FALSE)
```

